---
title: "Class05: Data, Viz, with GGPLOT"
author: "Bobbie Morales, A15443382"
format: pdf
---

Today we are playing with plotting and graphics in R.

There are lots of ways to make cool figures in R.
There is "base" R gaphics (`plot()`, `hist()` `boxplot()` etc)

There is also add on packages, like **ggplot**

```{r}
head(cars,3)
```
Let's plot this with "base" R: 

```{r}
plot(cars)
```

```{r}
head(mtcars)
```
#Let's plot `mpg` vs `disp`
```{r}
plot(mtcars$mpg, mtcars$disp, pch=10)
```

```{r}
hist(mtcars$mpg)
```

## GGPLOT

The main function in the GGPLOT 2 package is **ggplot()**. First I need to install the **ggplot2** package. I can install any package with the function `install.packages()`

```{r}
library(ggplot2)
ggplot(cars) +
  geom_point(aes(x = speed, y = dist))
```

> **N.B** I never want to run `install.package()` in quarto source document !!

```{r}
ggplot(cars) + 
  aes(speed) +
  geom_histogram()
```


Everyy ggplot needs at least 3 things: 

-The **data**(given with `ggplot(cars)`)
-The **aesthetic** mapping(given with `aes()`)
-The **geom** (given by `geom_point()`)

> For simple canned graphs "base" R is nearly always quicker 

### Adding more layers

Let's add a line

```{r}
ggplot(cars) +
  aes(x = speed, y = dist) +
  geom_point() + 
  geom_line() 
```

  
Let's add a line and a title, subtitle and caption as well as axis labels

```{r}
ggplot(cars) +
  aes(x = speed, y = dist) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE) + 
  labs(title = "Silly Plot", x = "Speed (MPH)",
       y = "Distance (ft)",
       caption = "when is tea time anyway??") + 
  theme_bw()
```


## Plot some expression data 

Read data file from online URL 

```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes)
```

> Q1. How many genes are in this wee dataset? 

There are `r nrow(genes)` in this dataset

```{r}
nrow(genes)
```
```{r}
colnames(genes)
ncol(genes)
```


>Q2. How many "up" regulated genes are there? 


```{r}
sum(genes$State == "up")
```

```{r}
round( table(genes$State)/nrow(genes))
```

There are `r sum(genes$State == "up")' UP genes

```{r}
table(genes$State)
```

```{r}
ggplot(genes) + 
geom_point(aes(x = Condition1, y = Condition2, col = Condition2))
```


```{r}
p <- ggplot(genes) + 
geom_point(aes(x = Condition1, y = Condition2, col = State)) + 
  scale_color_manual(values=c("blue", "grey", "red"))
```

```{r}
p + labs(title = "look at me")
```


Silly example of adding labels

```{r}
library(ggrepel)

 ggplot(genes) + 
  aes(x = Condition1, 
      y = Condition2,
      col = State, 
      label=Gene) + 
   geom_point() +
  scale_color_manual(values=c("blue", "grey", "red")) +
  geom_text_repel(max.overlaps = 5000) + 
   theme_bw()
```

##Going Further 
Playing with some different layers and the gapmider dataset..

```{r}
url <- "https://raw.githubusercontent.com/jennybc/gapminder/master/inst/extdata/gapminder.tsv"

gapminder <- read.delim(url)
```

```{r}
head(gapminder)
```

```{r}
tail(gapminder)
```

A first plot

```{r}
ggplot(gapminder) + 
  aes(x = gdpPercap,
      y = lifeExp, col=continent) + 
  geom_point(alpha = 0.2)  +
  facet_wrap(~continent) + 
  theme_bw()
```

```{r}
 p1 <- ggplot(gapminder) + 
  aes(x = gdpPercap,
      y = lifeExp, col=continent) + 
  geom_point(alpha = 0.2)  +
  facet_wrap(~continent) + 
  theme_bw()

p2 <- ggplot(cars) +
  aes(x = speed, y = dist) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE) + 
  labs(title = "Silly Plot", x = "Speed (MPH)",
       y = "Distance (ft)",
       caption = "when is tea time anyway??") + 
  theme_bw()
```
```{r}
library(patchwork)
(p1 + p2)
```
 
