---
title: "HW06"
author: "Bobbie Morales (A15443382)"
format: pdf
---

# Code to improve (A)

df <- data.frame(a=1:10, b=seq(200,400,length=10),c=11:20,d=NA)
df$a <- (df$a - min(df$a)) / (max(df$a) - min(df$a))
df$b <- (df$b - min(df$b)) / (max(df$b) - min(df$b))
df$c <- (df$c - min(df$c)) / (max(df$c) - min(df$c))
df$d <- (df$d - min(df$d)) / (max(df$a) - min(df$d))

# Create dataframe
```{r}
df <- data.frame(a=1:10, b=seq(200,400,length=10),c=11:20,d=NA)
```

# create function to normalize values in columm a to range btw 0-1 (make different variables comparable)

```{r}
normalize_formula <- function (x) {
  (x - min(x))/(max(x) - min(x))
  }
```

# Apply function only to numeric columns

```{r}
sapply(df[, sapply(df, is.numeric)], normalize_formula) 
```


# Code to improve (B)

library(bio3d)
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s1, chain="A", elety="CA")
s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")

# Create list to store kinases and function to read Protein Data Bank (PDB)

```{r}
library(bio3d)
kinase_list <-c("4AKE","1AKE","1E4Y")
structure_list <- lapply(kinase_list, read.pdb)
```

# Use trim function to create new, smaller PDB chain from kinases

```{r}
trim_pdb_function <- function(x) trim.pdb(x, chain="A", elety="CA")
trimmed_chains <- lapply(structure_list, trim_pdb_function)
```

# Extract B-factor values from atom data of trimmed chain

```{r}
extract_function <- function(x) {
  x$atom$b
}
b_factor_values <- lapply(trimmed_chains, extract_function)
```

# Plot B-factors vs Trimmed Chains
```{r}
plot_bfactors_values <- function(b_factors_values, trimmed_chains) {
  for (i in seq_along(b_factors_values)) {
    plotb3(b_factors_values[[i]], sse=trimmed_chains[[i]], typ="l", ylab="Bfactor")
  }
}
plot_bfactors_values(b_factor_values, trimmed_chains) 
```

